# Measurement Process Characterization

## Packages used in this chapter

```{r}
library(magrittr)
library(tidyverse)
```

## Characterization 

## Gauge R & R studies 

## Case Studies 

### Check standard 



#### Background and data 

>The measurements on the check standard duplicate certification measurements that were being made, during the same time period, on individual wafers from crystal #51939. For the check standard there were:

- J = 6 repetitions at the center of the wafer on each day
- K = 25 days

[Check standard for resistivity measurement](https://www.itl.nist.gov/div898/handbook/mpc/section6/mpc62.htm)

#### Reading the dataset

```{r}
check_standard <- read_table2("NIST data/MPC62_clean.DAT", col_names = TRUE) %>%
  rowid_to_column()
check_standard
```

#### Level-1 standard deviation

Measurements for *J* runs over *K* days for *L* runs are:

$$
Y_{lkj}(l = 1, \, \ldots, \, L, \,\, k=1, \,\ldots, \, K, \,\, j=1, \,\ldots, \, J)
$$
The level-1 repeatability (short term precision) is calcuated from the pooled standard deviation over days and runs

$$
s_{1lk} = \sqrt{\frac{1}{J-1} \sum_{j=1}^{J}{(Y_{lkj} - \overline{Y}_{lk{\small \, \bullet}})^2}}
$$

with

$$
\overline{Y}_{lk{\small \, \bullet}} = \frac{1}{J}\sum_{j=1}^{J}{\overline{Y}_{lkj}}
$$
As stated in the e-Handbook:
>An individual short-term standard deviation will not be a reliable estimate of precision if the degrees of freedom is less than ten, but the individual estimates can be pooled over the K days to obtain a more reliable estimate. 

The pooled level-1 standard deviation estimate with v = K(J - 1) degrees of freedom is
$$
s_1 = \sqrt{\frac{1}{K} \sum_{k=1}^{K} s_k^2}
$$

```{r}
s1_chkstd <- check_standard %>%
  mutate(Stdev_sq = Stdev^2) %$% 
  mean(Stdev_sq) %>% 
  sqrt()

s1_chkstd
```

Several comments on the code above. I've introduced the `%$%` operator. This allows me to use indivdual columns from my data frame and is useful for preforming mathematical operations on a specific column of data. It is from the **magrittr** package.

I find this type of code easy to read and understand. Describing the operations is simple, I'm just working from inside out of the equation:

- creating a new column of data that is $(Stdev)^2$
- finding the mean of that new column
- taking the square root of that number to give $s_1$.

#### Level-2 standard deviation (reproducibility)

$$
s_{chkstd} = {\large s}_2 = \sqrt{\frac{1}{K-1} \sum_{k=1}^{K} \left( \overline{Y}_{k \, \small{\bullet}} - \overline{Y}_{\small{\bullet}  \small{\bullet}} \right) ^2}
$$
with 

$$
\overline{Y}_{\small{\bullet} \small{\bullet}}  = \frac{1}{K} \sum_{k=1}^{K} \overline{Y}_{k \, \small{\bullet}}
$$

Which is simply the standard deviation of the daily measuremnts

```{r}
s2_chkstd <- check_standard %$%
  sd(Resistance)

s2_chkstd
```

#### Control chart for standard deviation - Precision

```{r}
UCL_precision_ckkstd <- s1_chkstd*sqrt(qf(0.95, 5, 125))
UCL_precision_ckkstd
```

```{r}
ggplot(check_standard) +
  geom_point(aes(rowid, Stdev)) +
  geom_hline(aes(yintercept = UCL_precision_ckkstd), colour = "red", linetype = "dashed") +
  labs(title =  "Precision control chart", subtitle = "Probe_ID 2362", x = "measurement", y = "ohm.cm", caption = "UCL calcuated at 95% level of confidence") +
  annotate("text", x = 0, y = 0.096, label = "UCL", colour = "red")
```

#### Control chart for measurement bias and variability

The control limits for monitoring the bias and long-term variability of resistivity with a Shewhart control chart are given by

$$
UCL=\text{Average} + 2 \cdot s_2 \\
Centerline=\text{Average} \\
LCL=\text{Average} − 2 \cdot s_2 \\
$$

```{r}
ggplot(check_standard) +
  geom_point(aes(rowid, Resistance)) +
  geom_hline(aes(yintercept = (mean(Resistance) + 2*s2_chkstd)), colour = "red", linetype = "dashed") +
  geom_hline(aes(yintercept = (mean(Resistance) - 2*s2_chkstd)), colour = "red", linetype = "dashed") +
  labs(title =  "Shewhart control chart", subtitle = "Probe_ID 2362", x = "measurement", y = "ohm.cm", caption = "Control limits calcuated with k = 2") +
  annotate("text", x = 0, y = 97.12, label = "UCL", colour = "red") +
  annotate("text", x = 0, y = 97.02, label = "LCL", colour = "red")
```


### Gauge study

#### Background and data
>Measurements on the check standards are used to estimate repeatability, day effect, and run effect	The effect of operator was not considered to be significant for this study; therefore, 'day' replaces 'operator' as a factor in the nested design. Averages and standard deviations from J = 6 measurements at the center of each wafer are shown in the table.

- J = 6 measurements at the center of the wafer per day
- K = 6 days (one operator) per repetition
- L = 2 runs (complete)
- Q = 5 wafers (check standards 138, 139, 140, 141, 142)
- R = 5 probes (1, 281, 283, 2062, 2362)

[Gauge study of resistivity probes](https://www.itl.nist.gov/div898/handbook/mpc/section6/mpc61.htm)


```{r}
gauge_study <- read_table2("NIST data/MPC61_clean.DAT", col_names = TRUE) %>%
  rowid_to_column()
gauge_study
```

#### Repeatability standard deviations

```{r}
ggplot(gauge_study) +
  geom_point(aes(as.factor(WAFER), STDDEV, colour = as.factor(DAY))) +
  facet_wrap(~ as.factor(PROBE), nrow = 1)
```

```{r}
ggplot(gauge_study) +
  geom_point(aes(as.factor(WAFER), STDDEV)) +
  facet_grid(as.factor(RUN) ~ as.factor(PROBE))
  
```

#### Effects of days and long-term stability

```{r}
ggplot(gauge_study) +
  geom_point(aes(rowid, AVERAGE, colour = as.factor(RUN))) +
  facet_grid(as.factor(WAFER) ~ as.factor(PROBE), scales = "free_y")
```


#### Differences among 5 probes

```{r}
probe_means_run <- gauge_study %>%
  group_by(PROBE, WAFER, RUN) %>%
  summarise(n = n(), probe_mean = mean(AVERAGE)) %>%
  unite(join_id, WAFER, RUN, sep = "_", remove = FALSE) %>%
  ungroup()

probe_means_run

wafer_means_run <- gauge_study %>%
  group_by(WAFER, RUN) %>%
  summarise(n = n(), wafer_means = mean(AVERAGE)) %>%
  unite(join_id, WAFER, RUN, sep = "_", remove = FALSE) %>%
  ungroup()

wafer_means_run

delta_probes <- left_join(probe_means_run, wafer_means_run, by = "join_id") %>%
  mutate(delta_probes_wafer = probe_mean - wafer_means)

delta_probes
```


```{r}
ggplot(delta_probes) +
  geom_line(aes(as.factor(WAFER.x), delta_probes_wafer, group = as.factor(PROBE), colour = as.factor(PROBE))) +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  facet_wrap(~ as.factor(RUN.x), ncol = 1) 
```

#### Analysis and interpretation

Table of estimates for probe #2362

>A graphical analysis shows repeatability standard deviations plotted by wafer and probe... The plots show that for both runs the precision of this probe is better than for the other probes.
>
>Probe #2362, because of its superior precision, was chosen as the tool for measuring all 100 ohm.cm resistivity wafers at NIST. Therefore, the remainder of the analysis focuses on this probe.

#### probe #2362

```{r}
probe_2362 <- gauge_study %>%
  filter(PROBE == 2362)

probe_2362
```

Pooled level-1 standard deviations (ohm.cm)

```{r}
s1_2362_1 <- probe_2362 %>%
  filter(RUN == 1) %>%
  mutate(Stdev_sq = STDDEV^2) %$% 
  mean(Stdev_sq) %>% 
  sqrt()

s1_2362_1

s1_2362_2 <- probe_2362 %>%
  filter(RUN == 2) %>%
  mutate(Stdev_sq = STDDEV^2) %$% 
  mean(Stdev_sq) %>% 
  sqrt()

s1_2362_2

s1_2362 <- probe_2362 %>%
  mutate(Stdev_sq = STDDEV^2) %$% 
  mean(Stdev_sq) %>% 
  sqrt()

s1_2362

```

Level-2 standard deviations (ohm.cm) for 5 wafers

```{r}
s2_2362 <- gauge_study %>%
  group_by(PROBE, WAFER, RUN) %>%
  filter(PROBE == 2362) %>%
  summarise(df = n()-1, probe_mean = mean(AVERAGE), probe_stdev = sd(AVERAGE), probe_stdev_sq = probe_stdev^2) %>%
  group_by(RUN) %>%
  summarise(s2_run = sqrt(mean(probe_stdev_sq)))

s2_2362

```

Over both runs
```{r}
s2_2352_all <- s2_2362 %>%
  mutate(s2_run_sq = s2_run^2) %$%
  mean(s2_run_sq) %>%
  sqrt()

s2_2352_all
  
```

```{r}
sd_2362_wafer <- gauge_study %>%
  group_by(PROBE, WAFER, RUN) %>%
  filter(PROBE == 2362) %>%
  summarise(probe_mean = mean(AVERAGE)) %>%
  mutate(
    run_number = case_when(
      RUN == 1 ~ "Run1",
      RUN == 2 ~ "Run2"
    )
  ) %>%
  select(PROBE, WAFER, probe_mean, run_number) %>% 
  group_by(WAFER) %>%
  summarise(sd_wafer = sd(probe_mean))


sd_2362_wafer

s3_2362 <- sd_2362_wafer %>%
  mutate(sd_wafer_sq = sd_wafer^2) %$%
  mean(sd_wafer_sq) %>%
  sqrt()
  
s3_2362
```

